The effects of binarity on planet occurrence rates measured by transit surveys

### 2017-12-11 sandbox/derivation.tex
propagate normalization fix until derivation end

For a given set of $(\delta_{\rm obs}, M\a)$  and $q$, the searchable distance for binaries is larger than singles by $\sqrt{1+L(qM\a)/L(M\a)}$. Thus,
\begin{equation}
	{N_d \over N_s^0}=\int_0^1 {n_b\over n_s}\left[1+{L(qM\a) \over L(M\a)}\right]^{3/2} f(q)\mathrm{d}q
	=\int_0^1 \frac{{\rm BF}}{1 - {\rm BF}}\left[1+{L(qM\a) \over L(M\a)}\right]^{3/2} f(q)\mathrm{d}q.
\end{equation}
If $\mathrm{BF}$ is independent of $q$, $L\sim M^\alpha$, and $f(q)\sim q^\beta$:
\begin{equation}
	{N_d \over N_s^0}
	=\frac{{\rm BF}}{1 - {\rm BF}}\cdot {1\over{1+\beta}}\int_0^1 \left(1+q^\alpha\right)^{3/2} q^\beta\mathrm{d}q,
\end{equation}
which may be written in a closed form using the hypergeometric function. 
Integrating by parts and using $u\equiv(1+q^\alpha)^{1/2}$, this reduces to
\begin{align}
	{N_d \over N_s^0}
	\notag
	&=\frac{{\rm BF}}{1 - {\rm BF}} \left(1\over{1+\beta}\right)^2
	\left\{
	\left[q^{\beta+1}\left(1+q^\alpha\right)^{3/2}\right]_0^1-{3\over2}\alpha \int_0^1 q^{\alpha+\beta}\left(1+q^\alpha\right)^{1/2} \mathrm{d}q \right\},\\
	&=\frac{{\rm BF}}{1 - {\rm BF}} \left(1\over{1+\beta}\right)^2
	\left[ 2^{3/2} - 3 \int_1^{\sqrt{2}} u^2(u^2-1)^{{1+\beta}\over\alpha}\mathrm{d}u \right].
\end{align}
For $\beta=0$, the last line is equivalent to Eq.(14).

@@

\edit1{MODIFIED UP TO THIS POINT}

\edit2{TODO: follow the above analytics to below limiting cases}

\subsubsection{Simplifying above general result}
To simplify, drop the explicit $M_a$ and BF 
dependences, and assume $R\propto M$, so $R(qM_a)/R(M_a) = q$.
Also, move all superscript numbers to subscripts.
Then the apparent rate density is
\begin{equation}
\Gamma\a(r\a) = \frac{1}{1+\mu}
\left(
    \Gamma_0(r\a) +
    \mu \left[
        \int \frac{{\rm d}q}{\mathcal{A}(q)} f(q) 
        \Gamma_1\left( \frac{r\a}{\mathcal{A}(q)} \right)
        +
        \int \frac{q {\rm d}q}{\mathcal{B}(q)} f(q)
        \Gamma_2\left(\frac{r\a}{\mathcal{B}(q)}\right) q^{2/3}
    \right]
\right).
\label{eq:simple_apparent_occ_rate}
\end{equation}
For the power law case when $L\sim M^\alpha$,
\begin{equation}
\mathcal{A}(q)
=(1+q^\alpha)^{-1/2},
\quad
\mathcal{B}(q)
=q^{-1}(1+q^{-\alpha})^{-1/2}
\label{eq:powerlaw_A_B}
\end{equation}

@@

We have $f(q)=\delta(q-1)$. The apparent rate density is
\begin{align}
\Gamma\a(r\a)
={1\over 1+\mu(\mathrm{BF})}
\left\{\Gamma_0(r\a)
+\mu(\mathrm{BF})\cdot \sqrt{2}\cdot\left[\Gamma_1(\sqrt{2}\,r\a)
+\Gamma_2(\sqrt{2}\,r\a)\right]
\right\},
\label{eq:apparent_rate_twin_binary}
\end{align}
where
\begin{equation}
	\mu(\mathrm{BF})=\int \mathrm{d}q\,f(q) 
	{\mathrm{BF}\over{1-\mathrm{BF}}}\left[1+{L(qM\a) \over 
	L(M\a)}\right]^{3/2}
	=2^{3/2}\cdot{\mathrm{BF}\over{1-\mathrm{BF}}}.
\end{equation}

@@

If $\Gamma_i(r, M)=Z_i\cdot\delta(r-r_0)$ (all the planets have the same 
radius), by using the identity $\delta(ax+b)=\delta(x+b/a)/|a|$,
\begin{align}
	\Gamma\a(r\a)
	={1 \over 1+\mu(\mathrm{BF})}
	\left[Z_0 \cdot \delta(r\a-r_0)+(Z_1+Z_2)\cdot 
	\mu(\mathrm{BF})\cdot 
	\delta\left(r\a-\frac{r_0}{\sqrt{2}}\right)\right].
    \label{eq:model1_apparent_rate}
\end{align}
This reproduces Eq.(9) of the draft (after removing the $p_{\rm det}$ 
terms)\footnote{
The above equation also follows by just thinking about the problem and 
explicitly writing out the answer (cf. LB's handwritten notes 
\texttt{2017/11/26.9}), though the formal derivation is a good check.
}.
In the limit of $\mathrm{BF}\to1$ ($\mu\to\infty$), the above formula yields 
$\Gamma\a=(Z^1+Z^2)\,\delta(r\a-{r_0/\sqrt{2}})$.

@@

Imagine instead that we took $\Gamma_i(r,M) = Z_i f(r) = Z_i 
r^\delta/\mathcal{N}_r$, for $f(r)$ the radius shape function, 
$\mathcal{N}_r$ the shape function's normalization, and $Z_i$ the number of 
planets per star.
From Eq.~\ref{eq:apparent_rate_twin_binary}, we get an apparent rate density
\begin{equation}
\Gamma_a(r_a) = \frac{r_a^\delta}{\mathcal{N}_r} \left[
\frac{Z_0}{1+\mu({\rm BF})}
+
2^{\frac{\delta+1}{2} } \frac{\mu({\rm BF})}{1+\mu({\rm BF})} \left(Z_1 + Z_2
\right)
\right].
\label{eq:model5_apparent_rate_density}
\end{equation}

@@

The stand-alone term ``true rate density'' is ill defined, and should probably 
be avoided.
In different contexts, it might refer to either a) 
$\Gamma_{\rm sel}(r)$, b) $\Gamma_{\rm vl}(r)$, or even c) $\Gamma_0(r)$, the 
rate around singles.
The meaning that most people probably {\it want} true rate density to have is
$\Gamma_0(r)$~---~simply ``I have a star that is single; how many planets are 
there?''.
This is also a sensible definition because it is {\it independent} of any 
complications of planet formation that would make $Z_1$ or $Z_2$ different 
from $Z_0$.
One way to deal with this, rather than making the arbitrary choice, would be 
to just plot out all the possible options.

@@

If we assume 
\begin{equation}
	L(M) \sim M^\alpha \sim R^\alpha,
\end{equation}
we find 
\begin{equation}
\mathcal{A}(q)
=(1+q^\alpha)^{-1/2},
\quad
\mathcal{B}(q)
=q^{-1}(1+q^{-\alpha})^{-1/2}.
\end{equation}

This gives a value of $\mu({\rm BF})$ stated in Eq.~\ref{eq:mu_final}.
Now recall Eq.~\ref{eq:simple_apparent_occ_rate} for the apparent rate 
density. We restate it here:
\begin{equation}
\Gamma\a(r\a) = \frac{1}{1+\mu}
\left(
\Gamma_0(r\a) +
\mu \left[
\int \frac{{\rm d}q}{\mathcal{A}(q)} f(q) 
\Gamma_1\left( \frac{r\a}{\mathcal{A}(q)} \right)
+
\int \frac{q {\rm d}q}{\mathcal{B}(q)} f(q)
\Gamma_2\left(\frac{r\a}{\mathcal{B}(q)}\right) q^{2/3}
\right]
\right).
\end{equation}

\subsubsection{Same Planets (Model \# 2)}
For our ``Model \#2'', we again want $\Gamma_i(r) = Z_i \cdot \delta(r-r_0)$.
To achieve this, write
\begin{equation}
f(q) = \frac{q^\beta (1+q^\alpha)^{3/2}}{\mathcal{N}_q},
\end{equation}
for $\mathcal{N}_q$ the normalization.
For $i=1$, call the integral on the left the $[\ldots]$ ``$I_1$''.
Then
\begin{align}
I_1(r_a) &\equiv \int \frac{{\rm d}q}{\mathcal{A}(q)} f(q) 
    \Gamma_1\left( \frac{r\a}{\mathcal{A}(q)} \right)
    \label{eq:defn_I1}
    \\
&= \frac{1}{\mathcal{N}_q} \int {\rm d}q\, q^\beta (1+q^\alpha)^2
    \Gamma_1(r_a (1+q^\alpha)^{1/2})
    \\
&= \frac{Z_1}{\mathcal{N}_q} \int {\rm d}q\, q^\beta (1+q^\alpha)^2
    \delta(r_a (1+q^\alpha)^{1/2} - r_0) \\
&= \frac{Z_1}{\mathcal{N}_q} \frac{2}{\alpha} \frac{r_0}{r_a^2} 
   \left(\left(\frac{r_0}{r_a}\right)^2 -1
       \right)^{\frac{1-\alpha}{\alpha}}
   \int {\rm d}q\, q^\beta (1+q^\alpha)^2 
       \delta\left( q - \left( \left(\frac{r_0}{r_a}\right)^2 -1
       \right)^{\frac{1}{\alpha}} \right) \\
&= \frac{Z_1}{\mathcal{N}_q} \frac{2}{\alpha} \frac{r_0}{r_a^2}
    \left( \left(\frac{r_0}{r_a}\right)^2 - 1  
    \right)^{\frac{\beta-\alpha+1}{\alpha}}
    \left( \frac{r_0}{r_a}\right)^4.
    \label{eq:model_2_i_1}
\end{align}
The penultimate line used the fact that $r_a(1+q^\alpha)^{1/2}=r_0$ has only 
one root:
\begin{equation}
q(r_a) = \left(\left(\frac{r_0}{r_a}\right)^2 - 1\right)^{1/\alpha},
\end{equation}
and for any function $f(q)$ with a single root $q_i$,
\begin{equation}
\delta(f(q)) = \frac{\delta(q-q_i)}{|f'(q_i)|}.
\end{equation}
Eq.~\ref{eq:model_2_i_1} is a good unit test for Monte Carlo numerics.

Attempting the same thing for $i=2$,
\begin{align}
I_2(r_a) &\equiv \int \frac{q {\rm d}q}{\mathcal{B}(q)} f(q)
\Gamma_2\left(\frac{r\a}{\mathcal{B}(q)}\right) q^{2/3}
\label{eq:defn_I2}
\\
&= \frac{Z_2}{\mathcal{N}_q} \int {\rm d}q\,
    q^{\beta+\frac{8}{3}} 
    (1+q^\alpha)^{3/2}
    (1+q^{-\alpha})^{1/2}
    \delta(r_a q (1+q^{-\alpha})^{1/2} - r_0).
\end{align}
Unfortunately, analytic roots of $r_a q (1+q^{-\alpha})^{1/2} = r_0$ do not 
exist.
If the roots are found numerically then
\begin{equation}
I_2(r_a) = \frac{Z_2}{\mathcal{N}_q}
    \sum_{q_i \in \{{\rm roots}\}}
q_i^{\beta+\frac{8}{3}} 
(1+q_i^\alpha)^{3/2}
(1+q_i^{-\alpha})^{1/2},
\end{equation}
but this doesn't really give much insight.

@@

Power law planets, a realistic mass ratio distribution.
In other words, now assume
\begin{equation}
    \Gamma_i(r,M) = Z_i f(r) = Z_i r^\delta / \mathcal{N}_r.
\end{equation}
Surprisingly, this might give more insight than Model \#2.

From Eq.~\ref{eq:simple_apparent_occ_rate}, we can write
\begin{equation}
\Gamma\a(r\a) = \frac{1}{1+\mu} \left(
\Gamma_0(r\a) + \mu  \left[
I_1(r_a) + I_2(r_a)
\right]
\right),
\end{equation}
where $I_1(r_a)$ and $I_2(r_a)$ are defined in the same way as in 
Eqs.~\ref{eq:defn_I1} and~\ref{eq:defn_I2}.
Doing the calculation, we find
\begin{equation}
I_1(r\a) = \frac{Z_1}{\mathcal{N}_q \mathcal{N}_r} r_a^\delta
\int {\rm d}q\,q^\beta (1+q^\alpha)^{\frac{\delta}{2}+2}
\end{equation}
and
\begin{equation}
I_2(r\a) = \frac{Z_2}{\mathcal{N}_q \mathcal{N}_r} r_a^\delta
\int {\rm d}q\,q^{\beta+\delta+\frac{8}{3}} 
               (1+q^\alpha)^{\frac{3}{2}}
               (1+q^{-\alpha})^{\frac{\delta+1}{2}}.
\end{equation}
The net effect of binarity is to just {\it scale} the overall distribution!

Note that the true rate density for the selected stars is
\begin{equation}
\Gamma_{\rm sel}(r) = \frac{r^\delta}{\mathcal{N}_r} \left[
\frac{Z_0}{1+2\mu} + \frac{\mu}{1+2\mu} (Z_1+Z_2)
\right].
\end{equation}
The correction factor with respect to selected stars, $X_\Gamma = 
\Gamma\a/\Gamma_{\rm sel}$, will be
\begin{align}
X_\Gamma &\equiv \left. \frac{\Gamma\a(r\a)}{\Gamma_{\rm sel}(r)} 
\right|_{r_a\rightarrow r,\ Z_i{\rm 's\ equal}} \\
&=
\frac{1+2\mu}{3(1+\mu)}
\left[1 + \mu
\left(
\int {\rm d}q\,q^\beta (1+q^\alpha)^{\frac{\delta}{2}+2} +
\int {\rm d}q\,q^{\beta+\delta+\frac{8}{3}} 
    (1+q^\alpha)^{\frac{3}{2}}
    (1+q^{-\alpha})^{\frac{\delta+1}{2}}
\right)\right].
\end{align}

For $\alpha = 3.5$, $\beta=0$, $\delta=-2.92$, the term in $(\ldots)\approx 
1.41425 \approx \sqrt{2}$, where the latter surprising numerical fluke holds 
to the first four decimal places.
For ${\rm BF}=0.44$, this yields $\Gamma\a/\Gamma_{\rm sel} = 1.28$, an 
apparent rate which is {\it higher} than the true rate by 28\%!
Note that this result is the same as $\Gamma\a/\Gamma0$ and 
$\Gamma\a/\Gamma_{\rm vl}$ in the limit where all the $Z_i$'s are equal.

>>>>

For a given set of $(\delta_{\rm obs}, M\a)$  and $q$, the searchable distance for binaries is larger than singles by $\sqrt{1+L(qM\a)/L(M\a)}$. Thus,
\begin{equation}
	{N_d \over N_s^0}=\int_0^1 {n_b\over n_s}\left[1+{L(qM\a) \over L(M\a)}\right]^{3/2} f(q)\mathrm{d}q
	=\int_0^1 \frac{{\rm BF}}{1 - {\rm BF}}\left[1+{L(qM\a) \over L(M\a)}\right]^{3/2} f(q)\mathrm{d}q.
\end{equation}
If $\mathrm{BF}$ is independent of $q$, $L\sim M^\alpha$, and $f(q)\sim q^\beta$:
\begin{equation}
	\mu = {N_d \over N_s^0}
	=\frac{{\rm BF}}{1 - {\rm BF}}\cdot {1\over{1+\beta}}\int_0^1 \left(1+q^\alpha\right)^{3/2} q^\beta\mathrm{d}q,
\end{equation}
which may be written in a closed form using the hypergeometric function. 
Integrating by parts and using $u\equiv(1+q^\alpha)^{1/2}$, this reduces to
\begin{align}
	\mu = {N_d \over N_s^0}
	\notag
	&=\frac{{\rm BF}}{1 - {\rm BF}} \left(1\over{1+\beta}\right)^2
	\left\{
	\left[q^{\beta+1}\left(1+q^\alpha\right)^{3/2}\right]_0^1-{3\over2}\alpha \int_0^1 q^{\alpha+\beta}\left(1+q^\alpha\right)^{1/2} \mathrm{d}q \right\},\\
	&=\frac{{\rm BF}}{1 - {\rm BF}} \left(1\over{1+\beta}\right)^2
	\left[ 2^{3/2} - 3 \int_1^{\sqrt{2}} u^2(u^2-1)^{{1+\beta}\over\alpha}\mathrm{d}u \right].
\end{align}
For $\beta=0$, the last line is equivalent to Eq.(14).

@@

\subsubsection{Simplifying above general result}
To simplify, drop the explicit BF dependences, and assume $R\propto M$, so 
$R(qM_a)/R(M_a) = q$. Also, move all superscript numbers to subscripts.
Then the apparent rate density is
\begin{align}
\notag
\Gamma\a(r\a,M\a) &= {1\over 1+\mu}\times
\left\{ \Gamma^0(r\a, M\a)+ 
\frac{{\rm BF}}{1 - {\rm BF}}
\left[ \int {\mathrm{d}q \over \mathcal{A}^4}\,f(q)\,\Gamma^1\left({r\a\over 
    \mathcal{A}}, 
M\a\right)\,
\right.   
\right. \\
& \quad\quad\quad\quad\quad \left.\left.
+\int {q \mathrm{d}q \over 
{\mathcal{A}^3\mathcal{B}}}\,f(q)\,\Gamma^2\left({r\a\over 
    \mathcal{B}}, 
qM\a\right)\,
q^{2/3} \right]	\right\}.  
\label{eq:simple_apparent_occ_rate}
\end{align}
For the power law case when $L\sim M^\alpha$,
\begin{equation}
\mathcal{A}(q)
=(1+q^\alpha)^{-1/2},
\quad
\mathcal{B}(q)
=q^{-1}(1+q^{-\alpha})^{-1/2}
\label{eq:powerlaw_A_B}
\end{equation}

@@

We have $f(q)=\delta(q-1)$. Since all stars have the exact same mass (both 
true mass and apparent mass), we will drop the $M\a$ dependence.
The apparent rate density is then
\begin{align}
\Gamma\a(r\a)
={1\over 1+\mu(\mathrm{BF})}
\left\{\Gamma_0(r\a)
+\mu(\mathrm{BF})\cdot \sqrt{2}\cdot\left[\Gamma_1(\sqrt{2}\,r\a)
+\Gamma_2(\sqrt{2}\,r\a)\right]
\right\},
\label{eq:apparent_rate_twin_binary}
\end{align}
where
\begin{equation}
	\mu(\mathrm{BF})
   	=\int_0^1 \frac{{\rm BF}}{1 - {\rm BF}}\left[1+{L(qM\a) \over 
    L(M\a)}\right]^{3/2} f(q)\,\mathrm{d}q
	=2^{3/2}\cdot{\mathrm{BF}\over{1-\mathrm{BF}}}.
\end{equation}

@@

If $\Gamma_i(r)=Z_i\cdot\delta(r-r_0)$ (all the planets have the same 
radius), by using the identity $\delta(ax+b)=\delta(x+b/a)/|a|$,
\begin{align}
	\Gamma\a(r\a)
	={1 \over 1+\mu(\mathrm{BF})}
	\left[Z_0 \cdot \delta(r\a-r_0)+(Z_1+Z_2)\cdot 
	\mu(\mathrm{BF})\cdot 
	\delta\left(r\a-\frac{r_0}{\sqrt{2}}\right)\right].
    \label{eq:model1_apparent_rate}
\end{align}
This reproduces Eq.(9) of the draft (after removing the $p_{\rm det}$ 
terms)\footnote{
The above equation also follows by just thinking about the problem and 
explicitly writing out the answer (cf. LB's handwritten notes 
\texttt{2017/11/26.9}), though the formal derivation is a good check.
}.
In the limit of $\mathrm{BF}\to1$ ($\mu\to\infty$), the above formula yields 
$\Gamma\a=(Z^1+Z^2)\,\delta(r\a-{r_0/\sqrt{2}})$.

@@

Imagine instead that we took $\Gamma_i(r) = Z_i f(r) = Z_i 
r^\delta/\mathcal{N}_r$, for $f(r)$ the radius shape function, 
$\mathcal{N}_r$ the shape function's normalization, and $Z_i$ the number of 
planets per star.
From Eq.~\ref{eq:apparent_rate_twin_binary}, we get an apparent rate density
\begin{equation}
\Gamma_a(r_a) = \frac{r_a^\delta}{\mathcal{N}_r} \left[
\frac{Z_0}{1+\mu({\rm BF})}
+
2^{\frac{\delta+1}{2} } \frac{\mu({\rm BF})}{1+\mu({\rm BF})} \left(Z_1 + Z_2
\right)
\right].
\label{eq:model5_apparent_rate_density}
\end{equation}

@@

The stand-alone term ``true rate density'' is ill defined, and should 
be avoided.
In different contexts, it might refer to either a) 
$\Gamma_{\rm sel}(r)$, b) $\Gamma_{\rm vl}(r)$, or even c) $\Gamma_0(r)$, the 
rate around singles.
The meaning that most people probably {\it want} true rate density to have is
$\Gamma_0(r)$~---~simply ``I have a star that is single; how many planets are 
there?''.
This is also a sensible definition because it is {\it independent} of any 
complications of planet formation that would make $Z_1$ or $Z_2$ different 
from $Z_0$.
One way to deal with this, rather than making the arbitrary choice, would be 
to just plot out all the possible options.
In general though, we will give highest preference to $\Gamma_0(r)$, since it 
can be compared with the results of other surveys.

@@

Now assume $f(q) = q^\beta/\mathcal{N}_q$, for $\mathcal{N}_q$ the 
normalization.
As previously noted, if we assume
\begin{equation}
	L(M) \sim M^\alpha \sim R^\alpha,
\end{equation}
we have 
\begin{equation}
\mathcal{A}(q)
=(1+q^\alpha)^{-1/2},
\quad
\mathcal{B}(q)
=q^{-1}(1+q^{-\alpha})^{-1/2}.
\end{equation}

This gives a value of $\mu({\rm BF})$ stated in Eq.~\ref{eq:mu_final}.
Now recall Eq.~\ref{eq:simple_apparent_occ_rate} for the apparent rate 
density. We restate it here:
\begin{align}
\notag
\Gamma\a(r\a,M\a) &= {1\over 1+\mu}\times
\left\{ \Gamma^0(r\a, M\a)+ 
\frac{{\rm BF}}{1 - {\rm BF}}
\left[ \int {\mathrm{d}q \over \mathcal{A}^4}\,f(q)\,\Gamma^1\left({r\a\over 
    \mathcal{A}}, 
M\a\right)\,
\right.   
\right. \\
& \quad\quad\quad\quad\quad \left.\left.
+\int {q \mathrm{d}q \over 
    {\mathcal{A}^3\mathcal{B}}}\,f(q)\,\Gamma^2\left({r\a\over 
    \mathcal{B}}, 
qM\a\right)\,
q^{2/3} \right]	\right\}.
\end{align}

Since in this model we will have stars with different masses, it makes more 
sense to keep the explicit ``$M\a$'' dependence in our equations.

\subsubsection{Same Planets (Model \# 2)}
For our ``Model \#2'', we assume
\begin{equation}
\Gamma_i(r,M) = Z_i \cdot \delta(r-r_0) M^\gamma / \mathcal{N}_M,
\end{equation}
where $\mathcal{N}_M$ has the correct units to make $Z_i$ dimensionless.
In most cases we'll take $\gamma=0$, but we'll carry it through the 
calculations here.

For $i=1$, call the integral on the left of the $[\ldots]$ ``$I_1$''.
Then
\begin{align}
I_1(r\a,M\a) &\equiv \int \frac{{\rm d}q}{\mathcal{A}^4} f(q) 
    \Gamma_1\left( \frac{r\a}{\mathcal{A}}, M\a \right)
    \label{eq:defn_I1}
    \\
&= \frac{1}{\mathcal{N}_q} \int {\rm d}q\, q^\beta (1+q^\alpha)^2
    \Gamma_1(r_a (1+q^\alpha)^{1/2}, M\a)
    \\
&= \frac{Z_1 M\a^\gamma}{\mathcal{N}_q \mathcal{N}_M} \int {\rm d}q\, q^\beta 
(1+q^\alpha)^2
    \delta(r_a (1+q^\alpha)^{1/2} - r_0) \\
&= \frac{Z_1 M\a^\gamma}{\mathcal{N}_q \mathcal{N}_M} \frac{2}{\alpha} 
\frac{r_0}{r_a^2} 
   \left(\left(\frac{r_0}{r_a}\right)^2 -1
       \right)^{\frac{1-\alpha}{\alpha}}
   \int {\rm d}q\, q^\beta (1+q^\alpha)^2 
       \delta\left( q - \left( \left(\frac{r_0}{r_a}\right)^2 -1
       \right)^{\frac{1}{\alpha}} \right) \\
&= \frac{Z_1 M\a^\gamma}{\mathcal{N}_q \mathcal{N}_M} \frac{2}{\alpha} 
\frac{r_0}{r_a^2}
    \left( \left(\frac{r_0}{r_a}\right)^2 - 1  
    \right)^{\frac{\beta-\alpha+1}{\alpha}}
    \left( \frac{r_0}{r_a}\right)^4.
    \label{eq:model_2_i_1}
\end{align}
The penultimate line used the fact that $r_a(1+q^\alpha)^{1/2}=r_0$ has only 
one root:
\begin{equation}
q(r_a) = \left(\left(\frac{r_0}{r_a}\right)^2 - 1\right)^{1/\alpha},
\end{equation}
and for any function $f(q)$ with a single root $q_i$,
\begin{equation}
\delta(f(q)) = \frac{\delta(q-q_i)}{|f'(q_i)|}.
\end{equation}
Eq.~\ref{eq:model_2_i_1} is a good unit test for numerics.

Attempting the same thing for $i=2$,
\begin{align}
I_2(r\a,M\a) &\equiv \int \frac{q {\rm d}q}{\mathcal{A}^3\mathcal{B}} f(q)
\Gamma_2\left(\frac{r\a}{\mathcal{B}}, qM\a\right) q^{2/3}
\label{eq:defn_I2}
\\
&=
Z_2 \int \frac{q {\rm d}q}{\mathcal{A}^3\mathcal{B}} f(q)
\delta\left(\frac{r\a}{\mathcal{B}} - r_0\right)
\frac{(qM\a)^{\gamma}}{\mathcal{N}_M}
q^{2/3}
\\
&= \frac{Z_2 M\a^\gamma}{\mathcal{N}_q\mathcal{N}_M} \int {\rm d}q\,
    q^{\beta+\gamma+\frac{8}{3}} 
    (1+q^\alpha)^{3/2}
    (1+q^{-\alpha})^{1/2}
    \delta(r_a q (1+q^{-\alpha})^{1/2} - r_0).
\end{align}
Unfortunately, analytic roots of $r_a q (1+q^{-\alpha})^{1/2} = r_0$ do not 
exist.
If the roots are found numerically then
\begin{equation}
I_2(r\a,M\a) = \frac{Z_2}{\mathcal{N}_q}
    \sum_{q_i \in \{{\rm roots}\}}
q_i^{\beta+\gamma+\frac{8}{3}} 
(1+q_i^\alpha)^{3/2}
(1+q_i^{-\alpha})^{1/2},
\end{equation}
but this doesn't really give much insight.

@@

Power law planets are a slightly more realistic mass ratio distribution.
If we now assume
\begin{equation}
    \Gamma_i(r,M) = Z_i \times \frac{r^\delta}{\mathcal{N}_r} \times
    \frac{M^\gamma}{\mathcal{N}_M},
\end{equation}
where $Z_i$ is dimensionless, and the normalization constants carry 
the units (each side has units $[r^{-1} M^{-1}]$).
From Eq.~\ref{eq:simple_apparent_occ_rate}, we can write
\begin{equation}
\Gamma\a(r\a,M\a) = \frac{1}{1+\mu} \left(
\Gamma_0(r\a,M\a) + \frac{{\rm BF}}{1-{\rm BF}} 
\left[ I_1(r_a,M\a) + I_2(r_a,M\a)
\right]
\right),
\end{equation}
where $I_1(r\a,M\a)$ and $I_2(r\a,M\a)$ are defined in the same way as in 
Eqs.~\ref{eq:defn_I1} and~\ref{eq:defn_I2}.
Doing the calculation, we find
\begin{align}
I_1(r\a,M\a) &= 
\int \frac{{\rm d}q}{\mathcal{A}} f(q)
\Gamma^1 \left(\frac{r\a}{\mathcal{A}},M\a\right)
\\
&=
\frac{Z_1}{\mathcal{N}_q \mathcal{N}_r \mathcal{N}_M} 
r\a^\delta M\a^\gamma
\int {\rm d}q\,q^\beta (1+q^\alpha)^{\frac{\delta+1}{2}}
\end{align}
and
\begin{align}
I_2(r\a,M\a) &=
\int \frac{q\,{\rm d}q}{\mathcal{A}^3 \mathcal{B}} f(q)
\Gamma^2 \left(\frac{r\a}{\mathcal{B}},qM\a\right)
\\
&= \frac{Z_2}{\mathcal{N}_q \mathcal{N}_r \mathcal{N}_M} 
r\a^\delta M\a^\gamma
\int {\rm d}q\,q^{\beta+\gamma+\delta+\frac{8}{3}} 
               (1+q^\alpha)^{\frac{3}{2}}
               (1+q^{-\alpha})^{\frac{\delta+1}{2}}.
\end{align}
The net effect of binarity is to just {\it scale} the overall distribution!

Note that the true rate density for a volume-limited sample is
\begin{align}
\Gamma_{\rm vl}(r,M) &= \frac{r^\delta M^\gamma}{\mathcal{N}_r\mathcal{N}_M} 
\left[
\frac{Z_0}{1+2(n_b/n_s)} + \frac{(n_b/n_s)}{1+2(n_b/n_s)} (Z_1+Z_2)
\right] \\
&= \frac{r^\delta M^\gamma Z_0}{\mathcal{N}_r\mathcal{N}_M}
\quad\quad ({\rm assuming}\ Z_0=Z_1=Z_2),
\end{align}
which when $Z_0=Z_1=Z_2$ is equal to $\Gamma_0$, and $\Gamma_{\rm sel}$.
The correction factor with respect to single stars, $X_\Gamma = 
\Gamma\a/\Gamma_0$, is then
\begin{align}
X_\Gamma &\equiv \left. \frac{\Gamma\a(r\a,M\a)}{\Gamma_0(r,M)} 
\right|_{r\a\rightarrow r,\ M\a\rightarrow M,\ Z_i{\rm 's\ equal}} \\
&=
\frac{1}{1+\mu}
\left[1 + \frac{2}{\mathcal{N}_q} \frac{{\rm BF}}{1-{\rm BF}}
\left(
\int {\rm d}q\,q^\beta (1+q^\alpha)^{\frac{\delta+1}{2}} +
\int {\rm d}q\,q^{\beta+\gamma+\delta+\frac{8}{3}} 
    (1+q^\alpha)^{\frac{3}{2}}
    (1+q^{-\alpha})^{\frac{\delta+1}{2}}
\right)\right].
\end{align}

For $\alpha = 3.5$, $\beta=0$, $\gamma=0$, $\delta=-2.92$, the sum of the 
integrals in $(\ldots)\approx 1.16393$. %171211_model_integrals.nb
For ${\rm BF}=0.44$, this yields $\Gamma\a/\Gamma_0 = 1.36$, an 
apparent rate which is {\it higher} than the true rate by 36\%!
Note that this result is the same as $\Gamma\a/\Gamma_{\rm sel}$ and 
$\Gamma\a/\Gamma_{\rm vl}$ in the limit where all the $Z_i$'s are equal.
